\racets: Faceted Execution in Racket

### 2018-09-11 paper/paper.tex
Adding changes for camera-ready

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros~\cite{Kohlbecker:1986, Dybvig:1992}. \racets provides facilities for
creating policies and faceting secure data with those
policies. \racets also extends several core forms in Racket to work
with faceted values (our implementation is detailed in
Section~\ref{sec:implementation}). We have used \racets to implement a
small server-based board-game (detailed in Sections~\ref{sec:overview}
and~\ref{sec:implementation}). Relevant related work is presented in
section~\ref{sec:related}. We see \racets as a promising prototype for
policy-agnostic programming in Racket, and conclude with discussion of
future directions in Section~\ref{sec:conclusion}.

@@

Although \code{mark-hit} will operate on sensitive data (the game
boards), it is written without any special machinery to maintain the
secrecy of \code{board}. Protecting data {w.r.t.} policies is instead handled
automatically and implicitly by a runtime monitor.
When Alice and Bob want to play a game, they both create a
\emph{label} to protect their data. A label is unique id mapped to a policy predicate
that takes a key (e.g., the current user's name) and returns true or false to indicate
permission for the label. Alice's label is used to annotate the data she wants
to be kept secret. Supposing Alice chooses to be player 1, she may use the
following label:

@@

@@

            \fcol{pc} & \fcol{\in} & \fcol{\text{PC}} & \fcol{=} & \fcol{\wp(\text{branch})} \\

@@

      \end{array}
    \end{displaymath}
    \\
    \hfill \textit{(Label Creation / Observation)}
    \\
    \[
    \begin{array}{cc}
      {\color{colorFAC}
        \inferrule[\textsc{Let-Label}]
          {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \langle \lambda x. ~ e, \rho' \rangle \\\\
           \alpha \not\in dom(\sigma') \and \sigma'' = \sigma' [\alpha \mapsto \langle \lambda x. ~ e, \rho' \rangle] \\\\
           e_2, \rho [ \ell \mapsto \alpha ], \sigma'' \Downarrow_{pc}^E \sigma''', v}
          {(\code{let-label} ~ \ell ~ e_1 ~ e_2), \rho, \sigma \Downarrow_{pc}^E \sigma''', v}}
          & 
          {\color{colorFAC}
          \inferrule[\textsc{Obs}]
                    {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \ell
                      \and e_2, \rho, \sigma' \Downarrow_{pc}^E \sigma'', v  \\\\
                      \langle (\lambda x.~e), \rho' \rangle = \sigma''(\ell) \\\\
                      e, \rho' [x \mapsto v], \sigma'' \Downarrow_{pc}^E \sigma''', v^{\pm} \\\\
                      e_3, \rho, \sigma''' \Downarrow_{pc}^E \sigma'''', v' \\\\
                      v'' = \fcol{\var{obs}(\ell, v', v^{\pm})}}
                    {(\code{obs} ~ e_1 ~ e_2 ~ e_3), \rho, \sigma \Downarrow_{pc}^E \sigma'''', v''}}
    \end{array}
    \]
    \\
    \hfill \textit{(Possibly-Faceted Application)} \boxed{\fcol{(v_1 ~ v_2), \rho, \sigma \Downarrow_{pc}^A \sigma, v}}
    \\
    \begingroup
    \color{colorMATH}
    \begin{displaymath}
      \begin{array}{c}
        \begin{array}{ccc}
          {\color{colorMATH}
          \inferrule[\textsc{App-$\star$}]
          { }
          {(\star ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma, \star}}
          & 
          {\color{colorMATH}
          \inferrule[\textsc{App-Base}]
          {e, \rho'[x \mapsto v], \sigma \Downarrow_{pc}^E \sigma', v' }
          {(\langle \lambda x.~e, \rho' \rangle ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{colorFAC}
          \inferrule[\textsc{App-Split}]
          {\{+\ell, -\ell\} \cap pc = \varnothing \and
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc \cup \{+\ell\}}^A \sigma', {v^+}' \\\\
            (v^- ~ v), \rho, \sigma' \Downarrow_{pc \cup \{-\ell\}}^A \sigma'', {v^-}' \and
            v' = \formfacet{\{+ \ell\}}{{v^+}'}{{v^-}'}}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma'', v'}}
        \end{array}
        \\
        \begin{array}{cc}
        \\\\
          {\color{colorFAC}
          \inferrule[\textsc{App-Facet-Pos}]
          {+\ell \in pc \\\\
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{colorFAC}
          \inferrule[\textsc{App-Facet-Neg}]
          {-\ell \in pc \\\\
            (v^- ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
        \end{array}
      \end{array}
    \end{displaymath}
    \endgroup
    \caption{Semantics of Faceted Execution}
    \label{fig:semantics}
\end{figure*}

@@

The rules \textsc{Box}, \textsc{Unbox}, and \textsc{Set} appear
similar to the standard implementation of boxes, but employ several
meta-functions to do their work. This is because box creation, reads,
and writes may occur within a privileged context, and care must be
taken to form facets when \rmth{pc} is nonempty. To understand why,
consider the following example\footnote{Our formal semantics elides
  \code{if}, though it may be obtained via a Church encoding if
  desired as in Austin et al.\cite{Austin:2012}. Our implementation
  includes \code{if} but not other constructs such as \code{cond}}:

@@

\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=other,numbers=none]
(facet l true (error ''this shouldn't get evaluated if <@\bmth{+\ell \in pc}@>''))
\end{lstlisting}

@@

\racets defines a \code{box} macro, along with \code{unbox} and
\code{set!}. We include the definition of \code{set!} in
Figure~\ref{fig:setbang}, which inlines the definition of the
\textit{write} metafunction from Section~\ref{sec:semantics} to
consider the case under which a facet is used when an address is
expected.

@@

\begin{figure*}
  \subfloat[]{%
    \includegraphics[height=1.1in]{a.png}%
    \label{fig:player1}%
  } 
  \quad
  \subfloat[]{%
    \includegraphics[height=1.1in]{b.png}%
    \label{fig:player1player2}%
  }
  \quad
  \subfloat[]{%
    \includegraphics[height=.7in]{c.png}%
    \label{fig:player2strike}%
  }
  \vspace{.3in}
  \quad
  \subfloat[]{%
    \includegraphics[height=1.1in]{d.png}%
    \label{fig:player1afterstrike}%
  }
  \caption{Screenshots from our Battleship case study.}
  \label{fig:case}
\end{figure*}

@@

\begin{lstlisting}[language=Racket,escapechar=|,name=case]
(define p1l (mkpol "player1"))
(define p1board
  (box (facet p1l
            (add-pieces (makeboard) '(1 2 ldots))
            (makeboard))))
\end{lstlisting}

@@

We plan to explore interactions with unfaceted code more in the
future, and believe it will an exciting direction. For example, once
execution escapes \racets, we have no guarantee that the privacy
policy won't be violated. One solution may be to implicitly perform an
\code{obs} based on the current \rmth{pc} at points where \racets
interacts with unfaceted modules. But we do not fully understand the
ramifications or ergonomics of this choice, and suspect there may be a
wide array of design choices to handle these module interactions
including security type systems and blame (to track which module
violated the privacy policy).

@@

Several other efforts into dynamic analysis for information flow are
worth noting. \citet{Stefan:2011} first presented \code{LIO}---a monad
(with implementation in Haskell) that tracks privilege of the current
program counter and forbids effects that would violate the security
policy. It may be surprising that \code{LIO}
works well for Haskell programs, given that faceted execution is
more precise than \code{LIO}---allowing values to become faceted
rather than halting the program. One key difference is that Haskell
programs emphasize purity while languages such as
JavaScript (the original target of faceted execution) do not, so much
of the machinery for faceted execution's effect on the store is less
interesting. Several authors have implemented related systems to
\code{LIO}, including variants of faceted execution \cite{Schmitz:2016}
and variants of \code{LIO} that extend its
power to arbitrary monad transformers~\cite{james}. We believe that it
would be possible to implement a variant of our technique that would
give similar insights to programs using \code{LIO}, though much of the
interesting machinery for handling state may be unnecessary.

>>>>

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros~\cite{Kohlbecker:1986, Dybvig:1992}. \racets provides facilities for
creating policies and faceting secure data with those
policies. \racets also rewrites several core forms in Racket to work
with faceted values (our implementation is detailed in
Section~\ref{sec:implementation}). For example, function application
must be treated specially to work with faceted values.
We have used \racets to implement a
small server-based board-game (detailed in Sections~\ref{sec:overview}
and~\ref{sec:implementation}). Relevant related work is presented in
section~\ref{sec:related}. We see \racets as a promising prototype for
policy-agnostic programming in Racket, and conclude with discussion of
future directions in Section~\ref{sec:conclusion}.

@@

Although \code{mark-hit} will operate on sensitive data (the game
boards), it is written without any special machinery to maintain the
secrecy of \code{board}. For example, nothing in \code{mark-hit}
prevents Bob from using \code{mark-hit} on Alice's board, but faceted
execution would ensure that if Bob tried to do this, he would always
see the result \code{false}. Protecting data {w.r.t.} policies is
instead handled automatically and implicitly by a runtime monitor.
When Alice and Bob want to play a game, they both create a
\emph{label} to protect their data. A label is unique id mapped to a
policy predicate that takes a key (e.g., the current user's name) and
returns true or false to indicate permission for the label. Alice's
label is used to annotate the data she wants to be kept
secret. Supposing Alice chooses to be player 1, she may use the
following label:

@@

            \fcol{\ell} & \fcol{\in} & \fcol{\text{label}} & \fcol{=} & \fcol{\ldots} \\

@@

            \fcol{pc} & \fcol{\in} & \fcol{\text{PC}} & \fcol{=} & \fcol{\{S \in \wp(\text{branch}) \mid \forall \ell. \{+\ell,-\ell\} \not\subseteq S\}} \\

@@

      \end{array}
    \end{displaymath}
    \\
    \hfill \textit{(Label Creation / Observation)}
    \\
    \[
    \begin{array}{cc}
      {\color{colorFAC}
        \inferrule[\textsc{Let-Label}]
          {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \langle \lambda x. ~ e, \rho' \rangle \\\\
           \alpha \not\in dom(\sigma') \and \sigma'' = \sigma' [\alpha \mapsto \langle \lambda x. ~ e, \rho' \rangle] \\\\
           e_2, \rho [ x \mapsto \alpha ], \sigma'' \Downarrow_{pc}^E \sigma''', v}
          {(\code{let-label} ~ x ~ e_1 ~ e_2), \rho, \sigma \Downarrow_{pc}^E \sigma''', v}}
          & 
          {\color{colorFAC}
          \inferrule[\textsc{Obs}]
                    {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \ell
                      \and e_2, \rho, \sigma' \Downarrow_{pc}^E \sigma'', v  \\\\
                      \langle (\lambda x.~e), \rho' \rangle = \sigma''(\ell) \\\\
                      e, \rho' [x \mapsto v], \sigma'' \Downarrow_{pc}^E \sigma''', v^{\pm} \\\\
                      e_3, \rho, \sigma''' \Downarrow_{pc}^E \sigma'''', v' \\\\
                      v'' = \fcol{\var{obs}(\ell, v', v^{\pm})}}
                    {(\code{obs} ~ e_1 ~ e_2 ~ e_3), \rho, \sigma \Downarrow_{pc}^E \sigma'''', v''}}
    \end{array}
    \]
    \\
    \hfill \textit{(Possibly-Faceted Application)} \boxed{\fcol{(v_1 ~ v_2), \rho, \sigma \Downarrow_{pc}^A \sigma, v}}
    \\
    \begingroup
    \color{colorMATH}
    \begin{displaymath}
      \begin{array}{c}
        \begin{array}{ccc}
          {\color{colorMATH}
          \inferrule[\textsc{App-$\star$}]
          { }
          {(\star ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma, \star}}
          & 
          {\color{colorMATH}
          \inferrule[\textsc{App-Base}]
          {e, \rho'[x \mapsto v], \sigma \Downarrow_{pc}^E \sigma', v' }
          {(\langle \lambda x.~e, \rho' \rangle ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{colorFAC}
          \inferrule[\textsc{App-Split}]
           {(v^+ ~ v), \rho, \sigma \Downarrow_{pc \cup \{+\ell\}}^A \sigma', {v^+}' \and
            {\{+\ell, -\ell\} \cap pc = \varnothing} \\\\
            (v^- ~ v), \rho, \sigma' \Downarrow_{pc \cup \{-\ell\}}^A \sigma'', {v^-}' \and
            v' = \formfacet{\{+ \ell\}}{{v^+}'}{{v^-}'}}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma'', v'}}
        \end{array}
        \\
        \begin{array}{cc}
        \\\\
          {\color{colorFAC}
          \inferrule[\textsc{App-Facet-Pos}]
          {+\ell \in pc \\\\
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{colorFAC}
          \inferrule[\textsc{App-Facet-Neg}]
          {-\ell \in pc \\\\
            (v^- ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
        \end{array}
      \end{array}
    \end{displaymath}
    \endgroup
    \caption{Semantics of Faceted Execution}
    \label{fig:semantics}
\end{figure*}

@@

The rules \textsc{Box}, \textsc{Unbox}, and \textsc{Set} appear
similar to the standard implementation of boxes, but employ several
meta-functions to do their work (As they are largely similar to those
in~\cite{Austin:2012}, we only detail $write$ here). This is because
box creation, reads, and writes may occur within a privileged context,
and care must be taken to form facets when \rmth{pc} is nonempty. To
understand why, consider the following example\footnote{Our formal
  semantics elides \code{if}, though it may be obtained via a Church
  encoding if desired as in Austin et al.\cite{Austin:2012}. Our
  implementation includes \code{if} but not other constructs such as
  \code{cond}}:

@@

\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=other,numbers=none]
(facet l true (error "this shouldn't get evaluated if <@\bmth{+\ell \in pc}@>"))
\end{lstlisting}

@@

\racets defines a \code{box} macro, along with \code{unbox} and
\code{set!}. We include the definition of \code{set!} in
Figure~\ref{fig:setbang}, which inlines the definition of the
\textit{write} metafunction from Section~\ref{sec:semantics} to
consider the case under which a facet is used when an address is
expected. In our current implementation, nothing prevents programmers
from misusing (or forgetting to use) our custom \code{box} and
\code{unbox} forms. We leave this to future work.

@@

\begin{figure*}
  \subfloat[]{%
    \includegraphics[height=2in]{a.png}%
    \label{fig:player1}%
  } 
  \quad
  \subfloat[]{%
    \includegraphics[height=2in]{b.png}%
    \label{fig:player1player2}%
  }
  \quad
  \subfloat[]{%
    \includegraphics[height=.7in]{c.png}%
    \label{fig:player2strike}%
  }
  \vspace{.3in}
  \quad
  \subfloat[]{%
    \includegraphics[height=2in]{d.png}%
    \label{fig:player1afterstrike}%
  }
  \caption{Screenshots from our Battleship case study.}
  \label{fig:case}
\end{figure*}

@@

\begin{lstlisting}[language=Racket,escapechar=|,name=case]
(define mkpol
  (lambda (player-name)
    (let-label l (lambda (x) (equal? x player-name)) l)))

(define p1l (mkpol "player1"))
(define p1board
  (box (facet p1l
            (add-pieces (makeboard) '(1 2 ldots))
            (makeboard))))
\end{lstlisting}

@@

We plan to explore interactions with unfaceted code more in the
future, and believe it will an exciting direction. For example, once
execution escapes \racets, we have no guarantee that the privacy
policy won't be violated. One solution may be to implicitly perform an
\code{obs} based on the current \rmth{pc} at points where \racets
interacts with unfaceted modules. But we do not fully understand the
ramifications or ergonomics of this choice, and suspect there may be a
wide array of design choices to handle these module interactions
including security type systems and blame (to track which module
violated the privacy policy). We speculate that we may be able to
piggy-back on top of Racket's existing contract system, and leave this
to future work.

@@

Several other efforts into dynamic analysis for information flow are
worth noting. \citet{Stefan:2011} first presented \code{LIO}---a monad
(with implementation in Haskell) that tracks privilege of the current
program counter and forbids effects that would violate the security
policy. It may be surprising that \code{LIO}
works well for Haskell programs, given that faceted execution is
more precise than \code{LIO}---allowing values to become faceted
rather than halting the program. One key difference is that Haskell
programs emphasize purity while languages such as
JavaScript (the original target of faceted execution) do not, so much
of the machinery for faceted execution's effect on the store is less
interesting. Several authors have implemented related systems to
\code{LIO}, including variants of faceted execution \cite{Schmitz:2016}
and variants of \code{LIO} that extend its
power to arbitrary monad transformers~\cite{james}.
